### 2024-08-31 article_codes.qmd
Add charts of WUI percentage with geofacet

# National WUI map

@@

# Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
#' @param .cache_path File path to save the results 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

@@

# get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximun distance in m (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

@@

  # Add colour table 
  coltab(wui) <- data.frame(id = 0:2, 
                            col = c("#E7EEE7", "#FF6633", "#003366"))

@@

  r <- .x
  # Identify levels (1 = Interface, 2 = Intermix)
  df_lev <- data.frame(id = 0:2, wui_class = c("No_wui", "Interface", "Intermix"))
  df_lev_lst <- replicate(nlyr(r), df_lev, simplify = FALSE)
  levels(r) <- df_lev_lst
  # Add colour table 
  df_col <- data.frame(id = 0:2, col = c("#E7EEE7", "#FF6633", "#003366"))
  df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)
  coltab(r) <- df_col_lst

@@

# Create local directory for caching big files
options(gisco_cache_dir = "C:/GISCO_spatial_data")

@@

dir.create("data/processed_data/tiles")
dir.create("data/processed_data/tiles/region")

@@

p1 <- ggplot() +
  geom_sf(data = world) +
  geom_sf(data = country_sf, fill = c("#D55E00", "#4E84C4")) +
  coord_sf(xlim = c(2500000, 5402285),
           ylim = c(1250000, 5500000),
           expand = FALSE) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "#E7EEE7")) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())

@@

plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(effis_fuel_cls, 
     col = effis_wild_veg_list_no$rgb_code,
     axes = FALSE, add = T)

@@

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(effis_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(effis_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

# Plot ----
plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(corine_fuel_veg, col = clc2018_wild_veg_list$rgb_code, 
     axes = FALSE, add = T)

@@

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(corine_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(corine_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

sr16_vrt <- terra::vrt(sr16r_path) 
levels(sr16_vrt) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
no_v_p <- project(no_v, "EPSG:25833")
plot(no_v_p, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(sr16_vrt, col = c("lightgreen", "darkgreen", "yellow"), axes = FALSE, add = T)

@@

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(sr16_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
plot(nmd_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(sr16_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"),
     axes = F, 
     legend = F,
     add = T)
plot(nmd_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"), 
     axes = F, 
     legend = F, 
     add = T)
title("B)", adj = 0)

@@

plot(region_r, col = "#E7EEE7", axes = FALSE, legend = FALSE)
plot(firEUrisk_fuel_250m, axes = FALSE, add = T)

@@

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(firEUrisk_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(firEUrisk_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

```{mermaid}
flowchart TD
  A[Grid 250m x 250m] --> B{"Building density > 6.17 #/km2"}
  B --  Yes  --> C{Vegetation > 50%}
  B --  No   --> D[No WUI]
  C --  Yes  --> E[Intermix] 
  C --  No   --> F{within 2.4 km of large forests (>5 km2)}
  F --  Yes  --> G[Interface]
  F --  No   --> H[No WUI]
  style E fill:#CC79A7
  style G fill:#F0E442
  style D fill:#D3D3D3
  style H fill:#D3D3D3
```

@@

3.  Create a raster indicating if the grid cell is within 2.4 km of a large (>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)

@@

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential, axes = F, legend = F, add = T)

@@

# plot_zoom <- function(.x) {
#   
#   zoom_area <- counties |> 
#     filter(NUTS_NAME == {{ .x }})
#   
#   wui_r_zoom <- wui_r |> 
#     crop(zoom_area) |> 
#     mask(zoom_area)
#   
#   # Add colour table 
#   coltab(wui_r_zoom) <- 
#     data.frame(id = 0:2, 
#                col = c("#E7EEE7", "#F0E442", "#CC79A7"))
#   
#   par(mfrow = c(1, 4))
#   
#   plot(wui_r_zoom$effis, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$effis))
#   # title( {{ .x }}, line = -4, adj = 0)
#   
#   plot(wui_r_zoom$corine, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$corine))
#   
#   plot(wui_r_zoom$firEUrisk,
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$firEUrisk))
#   
#   plot(wui_r_zoom$sr16,
#        axes = F, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$sr16)) 
#   
# }
# 
# # par(mfrow = c(1, 11))
# as.list(counties$NUTS_NAME) |> 
#   walk(plot_zoom)

@@

# Function for printing the table
tbl_summary_county <- function(.x) {
  .x |> 
    select(NUTS_NAME, name, starts_with("pct")) |> 
    gt(rowname_col = "name",
       groupname_col = "NUTS_NAME",
       row_group_as_column = FALSE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.background.color = "#E7EEE7",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) 
}

@@

@@

Area of wild vegetation in each county: Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution

```{r}
#| label: tbl-summary-table-areas
#| tbl-cap: Number of buildings and wildland vegetation percentage by county 
#| cache: true

# Number of buildings (use building_count?) ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

# Percentage vegetation county ----
# EFFIS
effis_county <- effis_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(effis_wild_veg))[1]/1000 * (res(effis_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_effis = area_km)

# CORINE
corine_county <- corine_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(corine_wild_veg))[1]/1000 * (res(corine_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_corine = area_km)

# National
sr16_county <- sr16_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(sr16_wild_veg))[1]/1000 * (res(sr16_wild_veg ))[2]/1000)
    ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  filter(CNTR_CODE  == "NO")

nmd_county <- nmd_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(nmd_wild_veg))[1]/1000 * (res(nmd_wild_veg ))[2]/1000)
         ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  filter(CNTR_CODE  == "SE")

@@

# FirEUrisk
firEUrisk_county <- firEUrisk_wild_veg |> 
  # Number of cells
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |>  
  st_as_sf() |> 
  # Area (number cells x area cells)
  mutate(
    area_km = wild_veg * ((res(firEUrisk_wild_veg))[1]/1000 * (res(firEUrisk_wild_veg ))[2]/1000)
         ) |> 
  # Get only values 
  st_drop_geometry() |> 
  select(NUTS_NAME, CNTR_CODE, area_km) |> 
  rename(area_km_firEUrisk = area_km)

# Joint tables
county_veg <- effis_county |> 
  left_join(corine_county) |> 
  left_join(national_county) |> 
  left_join(firEUrisk_county)

@@

# print table
df_summary |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |>
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN")) |> 
  gt(groupname_col = "CNTR_CODE",
     row_group_as_column = FALSE) |> 
  fmt_number(c(build_den, starts_with("perc")), decimals = 1) |> 
  fmt_number(c(sum, area), decimals = 0) |> 
  tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) |> 
  tab_spanner(
      label = md("**Area covered by vegetation [%]**"),
      columns = starts_with("perc")
    ) |> 
  tab_spanner(
      label = md("**Buildings**"),
      columns = c(sum, build_den)
    ) |> 
  cols_label(
      NUTS_NAME = md("**County**"),
      area = md("**Area [km2]**"),
      sum = md("**Count**"),
      build_den = md("**Density [#/km2]**"),
      perc_area_km_effis = md("**Effis**"),
      perc_area_km_corine = md("**Corine**"),
      perc_area_km_national = md("**National**"),
      perc_area_km_firEUrisk = md("**FirEurisk**")
    )

@@

Aggregate all maps in one figure (Note: data are re-sampled to 500859 cells in the plots)
```{r}
#| label: fig-wui-vegetation-type-distance
#| fig-cap: WUI by vegetation type map and threshold distance
#| cache: true

@@

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential_3, axes = F, legend = F, add = T)

@@

plot(region_r, col = "#E7EEE7", axes = F, legend = F)
plot(wui_potential_1, axes = F, legend = F, add = T)

@@

Plot all together
```{r}
#| label: fig-wui-threshold

p_1 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 1.00 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_3.08 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 3.08 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_6.17 <- ggplot() + 
  geom_spatraster(data = wui_r, maxcell = 2000000) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 6.17 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_1 / p_3.08 / p_6.17 +
  plot_layout(guides = "collect")

>>>>

# R scripts

@@

# Function: get_perc_veg_buffer ----
#' Function for calculating if the percentage of vegetation within a buffer area is higher than a threshold 
#' @description It returns a raster indicating if the percentage of flammable vegetation within a circumference, from the centre of the grid (buffer), higher than a threshold (YES = 1, NO = 0)
#' @param .veg raster layer indicating if the vegetation is considered as fuel or not (YES = 1, NO = 0). 
#' @param .radio_buffer Radio of the circumference.
#' @param .threshold Proportion: from 0-1 (default  = 0.5; i.e., 50%) 
get_perc_veg_buffer <- function(.veg,
                                .radio_buffer,
                                .threshold = 0.5){

@@

# get_dist_max ----
#' Get cells within a distance to vegetated area (YES --> 1; NO --> 0)  
#' @description It returns a raster with indicating the cells that are within the distance to the vegetated areas (YES = 1, NO = 0)
#' @param .x Raster with distances (from `terra::distance`) 
#' @param .max_distance Maximum distance in [m] (e.g., 2400 m)
get_dist_max <- function(.x, .max_dist) {

@@

  # Add colour table 
  coltab(wui) <- data.frame(id = 0:2, 
                            col = c("#FAFAFA", "#FF6633", "#003366"))

@@

  r <- .x
  # Identify levels (1 = Interface, 2 = Intermix)
  df_lev <- data.frame(id = 0:2, wui_class = c("No_wui", "Interface", "Intermix"))
  df_lev_lst <- replicate(nlyr(r), df_lev, simplify = FALSE)
  levels(r) <- df_lev_lst
  # Add colour table 
  df_col <- data.frame(id = 0:2, col = c("lightgrey", "#FF6633", "#003366"))
  df_col_lst <- replicate(nlyr(r), df_col, simplify = FALSE)
  coltab(r) <- df_col_lst

@@

# Create local directory for caching big files
dir.create("C:/GISCO_spatial_data")
options(gisco_cache_dir = "C:/GISCO_spatial_data")

@@

dir.create("data/processed_data")
dir.create("data/processed_data/tiles")
dir.create("data/processed_data/tiles/region")

@@

p1 <- ggplot() +
  geom_sf(data = world) +
  geom_sf(data = country_sf, fill = c("#D55E00", "#4E84C4")) +
  coord_sf(xlim = c(2500000, 5402285),
           ylim = c(1250000, 5500000),
           expand = FALSE) +
  theme_bw() +
  theme(panel.background = element_rect(fill = "lightgrey")) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank())

@@

plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(effis_fuel_cls, 
     col = effis_wild_veg_list_no$rgb_code,
     axes = FALSE, add = T)

@@

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(effis_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(effis_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

# Plot ----
plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(corine_fuel_veg, col = clc2018_wild_veg_list$rgb_code, 
     axes = FALSE, add = T)

@@

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(corine_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(corine_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

sr16_vrt <- terra::vrt(sr16r_path) 
levels(sr16_vrt) <- data.frame(id = 1:3, type =  c("spruce", "pine", "deciduous"))
no_v_p <- project(no_v, "EPSG:25833")
plot(no_v_p, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(sr16_vrt, col = c("lightgreen", "darkgreen", "yellow"), axes = FALSE, add = T)

@@

# Wildland vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(sr16_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
plot(nmd_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(sr16_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"),
     axes = F, 
     legend = F,
     add = T)
plot(nmd_wild_veg_large_dist_2400, 
     col = c(NA, "#0072B2"), 
     axes = F, 
     legend = F, 
     add = T)
title("B)", adj = 0)

@@

plot(region_r, col = "lightgrey", axes = FALSE, legend = FALSE)
plot(firEUrisk_fuel_250m, axes = FALSE, add = T)

@@

# Wildand vegetation (Yes = 1, No = NA)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(firEUrisk_wild_veg, col = "#009E73", axes = F, legend = F, add = T)
title("A)", adj = 0)
# Within a distance of 2.4 km (Yes = 1, No = 0)
plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(firEUrisk_wild_veg_large_dist_2400, col = c(NA, "#0072B2"), axes = F, legend = F, add = T)
title("B)", adj = 0)

@@

```{mermaid}
flowchart TD
    A["Grid 250m x 250m"] --> B{"Building density > 6.17 #/km2"}
    B -- Yes --> C{"Vegetation > 50%"}
    B -- No --> D["No WUI"]
    C -- Yes --> E["Intermix"]
    C -- No --> F{"Within 2.4 km of large 
    forest (>5 km)"}
    F -- Yes --> G["Interface"]
    F -- No --> H["No WUI"]
    style D fill:lightgrey
    style E fill:#003366,color:#FFFFFF
    style G fill:#FF6633
    style H fill:lightgrey

@@

3.  Create a raster indicating if the grid cell is within 2.4 km of a large (\>5 km2) fuel vegetation area (`wild_veg_large_dist_*`)

@@

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential, axes = F, legend = F, add = T)

@@

# plot_zoom <- function(.x) {
#   
#   zoom_area <- counties |> 
#     filter(NUTS_NAME == {{ .x }})
#   
#   wui_r_zoom <- wui_r |> 
#     crop(zoom_area) |> 
#     mask(zoom_area)
#   
#   # Add colour table 
#   coltab(wui_r_zoom) <- 
#     data.frame(id = 0:2, 
#                col = c("lightgrey", "#F0E442", "#CC79A7"))
#   
#   par(mfrow = c(1, 4))
#   
#   plot(wui_r_zoom$effis, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$effis))
#   # title( {{ .x }}, line = -4, adj = 0)
#   
#   plot(wui_r_zoom$corine, 
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0),
#        main =  names(wui_r_zoom$corine))
#   
#   plot(wui_r_zoom$firEUrisk,
#        axes = FALSE, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$firEUrisk))
#   
#   plot(wui_r_zoom$sr16,
#        axes = F, 
#        font.main = 1,
#        legend = FALSE,
#        mar = c(0, 0, 0, 0), 
#        main =  names(wui_r_zoom$sr16)) 
#   
# }
# 
# # par(mfrow = c(1, 11))
# as.list(counties$NUTS_NAME) |> 
#   walk(plot_zoom)

@@

# Function for printing the table
tbl_summary_county <- function(.x) {
  .x |> 
    select(NUTS_NAME, name, starts_with("pct")) |> 
    gt(rowname_col = "name",
       groupname_col = "NUTS_NAME",
       row_group_as_column = FALSE) |> 
    fmt_number(decimals = 1) |> 
    cols_label(
      name = "",
      pct_No_wui = md("**No WUI**"),
      pct_Interface = md("**Interface**"),
      pct_Intermix = md("**Intermix**"),
      pct_wui = md("**Total**")
    )  |> 
    tab_spanner(
      label = md("**WUI**"),
      columns = c(pct_Interface, pct_Intermix, pct_wui)
    ) |> 
    tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) 
}

@@

```
```{r}
#| label: fig-pct-wui-counties

# Add county ID
counties_cent_ID <- counties |>
  st_centroid() %>%
  mutate(x = st_coordinates(.)[,1],
         y = st_coordinates(.)[,2]) |> 
  arrange(desc(y), x) |> 
  mutate(ID = seq_along(counties$NUTS_NAME))

counties_ID <- counties |> 
  as_tibble() |> 
  left_join(counties_cent_ID, "NUTS_NAME") |> 
  st_sf()

# geofacet

mygrid <- data.frame(
  code = c("1 Troms og F.", 
           "2 Norrbotten", 
           "3 Nordland", 
           "4 Västerbotten", 
           "5 Trøndelag", 
           "7 Västernorrland", 
           "6 Jämtland", 
           "8 Møre og R.", 
           "10 Innlandet", 
           "11 Dalarna", 
           "9 Gävleborg", 
           "12 Vestland",
           "15 Oslo",
           "14 Viken",
           "13 Uppsala",
           "16 Västmanland",
           "20 Örebro", 
           "17 Värmland",
           "22 Rogaland", 
           "19 Vestfold og T.",
           "18 Stockholm", 
           "21 Södermanland",
           "24 Östergötland",
           "25 Västra Götaland",
           "23 Agder", 
           "29 Halland", 
           "27 Jönköping", 
           "28 Kalmar", 
           "26 Gotland", 
           "30 Kronoberg", 
           "31 Blekinge",
           "32 Skåne"),
  row = c(2, 2, 3, 3, 4, 4, 4, 5, 5, 5,
          5, 6, 6, 6, 6, 6, 6, 6, 7, 7, 
          7, 7, 7, 7, 8, 8, 8, 8, 9, 9, 9, 10),
  col = c(6, 7, 6, 7, 5, 7, 6, 4, 5, 6,
          7, 3, 5, 4, 9, 8, 7, 6, 3, 4, 
          9, 8, 7, 6, 4, 6, 7, 8, 9, 6, 7, 6),
  name = c("Troms og Finnmark",
           "Norrbotten", 
           "Nordland",
           "Västerbotten", 
           "Trøndelag", 
           "Västernorrland",
           "Jämtland", 
           "Møre og Romsdal",
           "Innlandet", 
           "Dalarna", 
           "Gävleborg",
           "Vestland", 
           "Oslo", 
           "Viken", 
           "Uppsala",
           "Västmanland",
           "Örebro", 
           "Värmland",
           "Rogaland", 
           "Vestfold og Telemark",
           "Stockholm", 
           "Södermanland", 
           "Östergötland", 
           "Västra Götaland", 
           "Agder", 
           "Halland", 
           "Jönköping", 
           "Kalmar",
           "Gotland", 
           "Kronoberg", 
           "Blekinge", 
           "Skåne"),
  stringsAsFactors = FALSE
)

# Maps

p_map <- ggplot() +
  geom_sf(data = counties_ID, 
          fill = "#FAFAFA",
          col = "lightgrey",
          size = 0.5) +
  geom_sf_text(data = counties_ID, 
               aes(label = ID),
               col = "red",
               size = 3) +
  theme_void()

p_bar <- counties_wui |> 
  select(NUTS_NAME, name, starts_with("pct")) |> 
  select(-pct_wui) |> 
  pivot_longer(cols = starts_with("pct"), 
               names_to = "type",
               values_to = "value") |> 
  mutate(type = str_remove_all(type, "pct_"),
         NUTS_NAME = str_remove_all(NUTS_NAME, "s län"),
         NUTS_NAME = str_remove_all(NUTS_NAME, " län")) |> 
  mutate(name = case_when(name == "corine" ~ "C",
                          name == "effis" ~ "E",
                          name == "national" ~ "N",
                          name == "firEUrisk" ~ "F")) |>
  mutate(name = factor(name, levels = c("N", "C", "E", "F"))) |> 
  mutate(type = factor(type, levels = c("No_wui", "Interface", "Intermix"))) |> 
  ggplot() +
  geom_bar(aes(x = name,
               y = value,
               fill = type),
           col = "grey",
           stat = "identity",
           position = "fill"
           ) +
  scale_fill_manual(values = c("#FAFAFA", "#FF6633", "#003366")) +
  geofacet::facet_geo(~NUTS_NAME,
                      grid = mygrid, 
                      label = "code"
                      ) +
  labs(x = "",
       y = "") +
  theme_bw() +
  theme(
    strip.text = element_text(size = 9, face = "bold")
  )

ggsave(plot = p_map, 
       filename = "figures/counties_map.png", 
       width = 3, height = 3)

ggsave(plot = p_bar, 
       filename = "figures/counties_wui_perc.png", 
       width = 15, height = 20)

@@

![](figures/counties_map.png)
![WUI percentage by county](figures/counties_wui_perc.png)

Area of wild vegetation by county (@tbl-summary-table-areas): Number vegetation cells (centroid is inside the polygon), sum then and calculate the area based on the raster resolution.

```{r}
#| label: tbl-summary-table-areas
#| tbl-cap: Number of buildings and wildland vegetation percentage by county (and total)
#| cache: true

# Number of buildings by county  ----
county_build_count <- build_counts |> 
  # Number of buildings per county
  extract(y = vect(counties),
          fun = "sum",
          na.rm = TRUE,
          bind = TRUE) |> 
  # sf object
  st_as_sf() %>% 
  # Area polygons
  mutate(area = units::set_units(st_area(.), "km2"))  |> 
  # Building density by county
  mutate(build_den = sum/area)

# Percentage vegetation by county ----

# Function for getting vegetation area by counties 
f_pct_veg <- function(df) {
  df |> # sf object with vegetation data  
    # Number of cells
    extract(y = vect(counties),
            fun = "sum",
            na.rm = TRUE,
            bind = TRUE) |>  
    st_as_sf() |> 
    # Area (number cells x area cells)
    mutate(
      area_km = wild_veg * ((res(df))[1]/1000 * (res(df ))[2]/1000)
    ) |> 
    # Get only values 
    st_drop_geometry() |> 
    select(NUTS_NAME, CNTR_CODE, area_km)
}

## Run for each vegetation type

# Effis
effis_county  <- f_pct_veg(effis_wild_veg) |> 
    rename(area_km_effis = area_km)

# Corine
corine_county <- f_pct_veg(corine_wild_veg) |> 
  rename(area_km_corine = area_km)

# National
sr16_county <- f_pct_veg(sr16_wild_veg) |> 
  filter(CNTR_CODE  == "NO")
nmd_county  <- f_pct_veg(nmd_wild_veg)  |> 
  filter(CNTR_CODE  == "SE")

@@

# FirEUrisk
firEUrisk_county <- f_pct_veg(firEUrisk_wild_veg) |> 
  rename(area_km_firEUrisk = area_km)

# Joint tables
county_veg <- national_county |> 
  left_join(corine_county) |> 
  left_join(effis_county) |> 
  left_join(firEUrisk_county)

@@

# Summary tables by country and county ----
summary_country <- df_summary |> 
  select(CNTR_CODE, sum, area, starts_with("area_km")) |> 
  group_by(CNTR_CODE) |> 
  summarise_all(sum) |> 
  mutate(NUTS_NAME = "Total",
         build_den = sum/area) |> 
  mutate(across(.cols = starts_with("area_km"),
                .fns = ~ 100 * (.x / as.numeric(area)),
                .names = "perc_{.col}")) |> 
  relocate(build_den, .after = area) |> 
  relocate(NUTS_NAME, .after = CNTR_CODE) |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |> 
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN"))

summary_county <- df_summary |> 
  select(CNTR_CODE, NUTS_NAME, area,
         sum, build_den,
         starts_with("perc")) |>
  arrange(CNTR_CODE, NUTS_NAME) |> 
  mutate(area = as.numeric(area),
         build_den = as.numeric(build_den),
         CNTR_CODE = case_when(CNTR_CODE == "NO" ~ "NORWAY",
                               CNTR_CODE == "SE" ~ "SWEDEN"))

# Print table ----
summary_county |> 
  rbind(summary_country) |> 
  gt(groupname_col = "CNTR_CODE",
     row_group_as_column = FALSE) |> 
  fmt_number(c(build_den, starts_with("perc")), decimals = 1) |> 
  fmt_number(c(sum, area), decimals = 0) |> 
  tab_options(
      table.font.size = 9,
      row_group.background.color = "lightgrey",
      row_group.padding = px(1),
      row_group.font.weight = "bold",
      data_row.padding = px(1)
    ) |> 
  tab_spanner(
      label = md("**Area covered by vegetation [%]**"),
      columns = starts_with("perc")
    ) |> 
  tab_spanner(
      label = md("**Buildings**"),
      columns = c(sum, build_den)
    ) |> 
  cols_label(
      NUTS_NAME = md("**County**"),
      area = md("**Area [km2]**"),
      sum = md("**Count**"),
      build_den = md("**Density [#/km2]**"),
      perc_area_km_national = md("**National**"),
      perc_area_km_effis = md("**Effis**"),
      perc_area_km_corine = md("**Corine**"),
      perc_area_km_firEUrisk = md("**FirEurisk**")
    ) |> 
  tab_style(
    style = cell_text(weight = "bold"),  # Apply bold text
    locations = cells_body(
      rows = c(33, 34)  # specify the row number
    )
  )

@@

Aggregate all maps in one figure (Note: data are re-sampled to 500859 cells in the plots)

```{r}
#| label: fig-wui-vegetation-type-distance
#| fig-cap: WUI by vegetation type map and threshold distance
#| fig-height: 15
#| cache: true

@@

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential_3, axes = F, legend = F, add = T)

@@

plot(region_r, col = "lightgrey", axes = F, legend = F)
plot(wui_potential_1, axes = F, legend = F, add = T)

@@

Plot all together

```{r}
#| label: fig-wui-threshold
#| fig-height: 15

p_1 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 1.00 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_3.08 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 3.08 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

p_6.17 <- ggplot() + 
  geom_spatraster(data = wui_r) +
  facet_wrap(~lyr, nrow = 1) +
  labs(title = "d = 6.17 #/km2") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

# NOTE: geom_spatraster change the resolution of the map
p_1 / p_3.08 / p_6.17 +
  plot_layout(guides = "collect")
